---
title: "Project 1 DATA 607 Jun Pan"
author: "Jun Pan"
date: "September 21, 2018"
output: html_document
---

#Project 1 Chess Rating
In this project, you're given a text file with chess tournament results where the information has some structure.  Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players: Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents.

Install packages for this project
```{r}
if (!require('DT')) install.packages('DT')
if (!require('stringr')) install.packages('stringr')
if (!require('ggplot2')) install.packages('ggplot2')
```

```{r}
library(DT)
library(stringr)
library(ggplot2)
library(dplyr)
library(readr)
library(tidyr)
library(knitr)
```


load chess tournament results
```{r}
raw = readLines("C:/Users/tbao/Desktop/CUNY MSDS notes/607/project 1 wk4/tournamentinfo.txt")
raw = str_replace_all(raw,"-","") 
print(raw[1:9])
```


Data groping and splitting
Find the table breaks
```{r}
b0 <- 0
b1 <- unname(str_locate_all(pattern = '\\|', raw[5])[[1]][1,1])
b2 <- unname(str_locate_all(pattern = '\\|', raw[5])[[1]][2,1])
b3 <- unname(str_locate_all(pattern = '\\|', raw[5])[[1]][3,1])
b4 <- max(nchar(raw))
```

Data seperating
```{r}
# Group1 = Num, Name, Points, Opponent IDs
g1row <- seq(5, 196, 3)
# Group2 = State, Rating
g2row <- seq(6, 196, 3)
# subset for easy searching
group1 <- raw[g1row]
group2 <- raw[g2row]
```

Data Field Creation and Extraction
Player Name
```{r}
namesub <- substr(group1, b1+1, b2-2)
namesub <- str_trim(namesub)
PlayerName <- str_to_title(namesub)
head(PlayerName)
```

Player State
```{r}
statesub <- substr(group2, b0, b1-1)
State <- str_trim(statesub)
head(State)
```

```{r}
# dataframe
chess <- data.frame(PlayerName, State)
```


Total Points
```{r}
pointsub <- substr(group1, b2+1, b3-1)
head(pointsub)
```

```{r}
chess$TotalPoints <- as.numeric(unlist(str_extract_all(raw, "(\\d)\\.(\\d)")))
```

Player Pre-Rating
```{r}
presub <- substr(group2, b1+1, b2-1)
presub <- str_extract(presub, ': *\\d{2,}')
head(presub)
```

```{r}
chess$PreRating <- as.integer(str_extract(presub, '\\d{2,}'))
```

Avg Opponent Pre-Rating
```{r}
oppsub <- substr(group1, b3+1, b4)
oppsub <- str_extract_all(oppsub, '\\b\\d{1,}')
oppsub <- as.matrix(oppsub)

calculation <- function(l, p ){
  temp <- l[p]
    for (place in temp){
        rating <- 0
        counter <- 0
        for(i in place) {
            counter <- counter + 1
            rating <- rating + chess$PreRating[as.numeric(i)]
        }
        rating <- round(rating / counter)
    }
  return (rating)
}

chess$AvgOppPreRating <- apply(oppsub, 1, calculation)
```

Review Final Dataset
```{r}
datatable(chess)
```

Export in .csv
```{r}
write.csv(chess, "chessData.csv", row.names=FALSE)
```

Data visualization
```{r}
x <- ggplot(chess, aes(PreRating, AvgOppPreRating)) + geom_point(aes(color=TotalPoints)) + ggtitle("Pre-Rating VS Avg Opponent Pre-Rating by Total Points Gained")
x
```


